Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-17 nips_2018.tex
improved big table

, the program synthesis and AI communities have had many successes,
not just for text editing~\cite{gulwani2011automating} but also for graphics
programs~\cite{ellis2017learning}, planning
procedures~\cite{devlin2017neural}, and many others.  However, the
success of these systems hinges upon having a carefully
hand-engineered \textbf{Domain Specific Language (DSL)}.  DSLs
impart prior knowledge of a domain
by providing a restricted set
of finely-tuned programming primitives: for
%that are finely tuned to the problems in the domain: for
text editing,
these are primitives like appending and splitting on characters.
In this work,
we consider the problem of
building agents that solve program learning tasks,
and also the problem of acquiring
the prior knowledge necessary to quickly
solve these tasks (Figure~\ref{firstPageFigure}).
Our solution is an algorithm that learns a DSL
while jointly training a neural network to help write programs in the learned DSL.
\begin{figure}[h]
  \begin{minipage}[c]{0.68\textwidth}
\hspace{-0.5cm}      \begin{tabular}{r|c|}\cmidrule{2-2}
  \sc{Task}&
  \begin{tabular}{rcl}
  Nancy FreeHafer&$\longrightarrow$& Dr. Nancy\\
% Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
 Jane Goodall&$\longrightarrow$&???
  \end{tabular}\\\cmidrule{2-2}
  \sc{Program}&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_1$\code{ s ' '))}\\\cmidrule{2-2}
    \sc{DSL}
&\begin{tabular}{l}
    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b ($\lambda$ (x y)}\\
    \phantom{$f_0($\code{a}$,$\code{b}$) \,=\, $\code{(}}\code{(cons x y)))}\\
    \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s ($\lambda$ (x a)}\\
    \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}\\\cmidrule{2-2}
  \end{tabular}
  \end{minipage}%
  \begin{minipage}[c]{0.31\textwidth}
      \caption{\textbf{Task}: Few-shot learning problem. Model solves tasks by writing \textbf{program}s, and jointly learns a library of reusable subroutines that are shared across multiple tasks, called a \textbf{Domain Specific Language (DSL)}. Program writing is guided by a neural network trained jointly with the library.}\label{firstPageFigure}
    \end{minipage}
\end{figure}

@@

We apply \system to four domains:
list processing; FlashFill-style~\cite{gulwani2011automating} string editing; and symbolic regression.
 For each of these we initially provide a generic
 set of programming primitives.
Our algorithm then discovers
its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{figure}
  \begin{tabular}{lcl}
    \toprule
    Domain&Example Task&Part of the learned DSL\\\midrule
    Lists&
    \begin{tabular}{rcl}
      \code{[7\, 0\, 2]}&$\longrightarrow$&\code{[7\, 0\, 2\, 4]} \\
      \code{[3\, 9]}&$\longrightarrow$&\code{[3\, 9\, 4]}
    \end{tabular}
    &
    \begin{tabular}{l}
      \code{(foldr nil ($\lambda$ (a b) (cons a b)))}\\
          \hspace{0.5cm} \emph{(appends lists)}
      \end{tabular}

    \\    Strings&
    \begin{tabular}{rcl}
      Temple Anna H& $\longrightarrow$& TAH\\
      Lara Gregori&$\longrightarrow$&LG
      \end{tabular}
    &
    \begin{tabular}{l}
      \code{(map ($\lambda$ (x) (if (= x a) b x)))}\\
      \hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
      \end{tabular}

    \\Regression&\includegraphics[width = 1cm]{figures/functions/476.png}&\begin{tabular}{l}
      \code{(+ (* real x) real)} \\
          \hspace{0.5cm} \emph{(a linear function of x)}
      \end{tabular}

    \\\bottomrule
  \end{tabular}
  \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
%  \vspace{-4em}
\end{figure}

\begin{figure}[hb]
\makebox[\textwidth][c]{
\scriptsize
\renewcommand\code\texttt
\renewcommand\codechar[1]{\texttt{"#1"}}
\begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}}\toprule
  Lists&Strings&Symbolic Regression\\\midrule
  \begin{tabular}{c}
      % \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
      % \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
      % \code{$f(x) = $}\code{($f_0$ nil $x$)}\\\\
      \code{[99\, 4\, 2]}$\to $\code{99}\\
      \code{[2\, 1\, 4\, 6]}$\to $\code{6}\\
      \code{[8\, 7\, 7\, 7]}$\to $\code{8}\\
      \code{$f(\ell) = $}\code{($f_0$ $\ell$)}\\\\
      \code{[7\, 3\, 14\, 6\, 9]}$\to $\code{false}\\
      \code{[9\, 4\, 3\, 4]}$\to $\code{true}\\
      \code{$f(\ell) = $}\code{($f_1$ $\ell$ k)}
  \end{tabular}&
  \begin{tabular}{c}
    +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
    +83 973-831$\to $83.973.831\\
    $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} ($f_0$ \codechar{.} \codechar{ } (cdr s)))}\\\\
Temple Anna H $\to $TAH\\
Lara Gregori$\to $LG\\
$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}\\

@@

  \begin{tabular}{c}
    \includegraphics[width = 3em]{figures/functions/4.png}\\
    \code{$f($x$) = $($f_1$ -0.8 1.6)}\\
    \includegraphics[width = 3em]{figures/functions/112.png}\\
    \code{$f($x$) = $($f_0$ -0.8 1.6)}\\

@@

  \begin{tabular}{l}
    % $f_0($\code{r}$,\ell) \,=\, $\code{(fold $\ell$ r ($\lambda$ (x a)}\\
    % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(fold}}\code{(cons (index (length a) $\ell$) a)))}\\
      % \hspace{0.5cm}($f_1$: \emph{Get the largest number})\\
    $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a) (if (> a x) a x)))}\\
      \hspace{0.5cm}($f_1$: \emph{Get the largest number in $\ell$})\\
    $f_1(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (empty$?$ $\ell$)}\\
    \phantom{$f_1(\ell,$\code{k}$) \,=\, $\code{(foldr}}
    \code{($\lambda$ (x a) (if a a (= k x))))}\\
      \hspace{0.5cm}($f_1$: \emph{Whether $\ell$ contains $k$})\\
  \end{tabular}&

  \begin{tabular}{l}
          $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x) (if (= x a) b x)) s)}\\
      \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (cdr (if (= c x) s a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s empty$?$ car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})\\
    \end{tabular}&

  \begin{tabular}{l}
    $f_0($\code{x}$)\,=\,$\code{(+ x real)}\\
  $f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} \\
  $f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}\\
    \end{tabular}

@@

\end{tabular}}
\end{figure}

>>>>

, the program synthesis and AI communities have had many successes,
not just for text editing~\cite{gulwani2011automating} but also for graphics
programs~\cite{ellis2017learning}, planning
procedures~\cite{devlin2017neural}, and many others.  However, the
success of these systems hinges upon having a carefully
hand-engineered \textbf{Domain Specific Language (DSL)}.  DSLs
impart prior knowledge of a domain
by providing a restricted set
of finely-tuned programming primitives: for
%that are finely tuned to the problems in the domain: for
text editing,
these are primitives like appending and splitting on characters.
In this work,
we consider the problem of
building agents that solve program learning tasks,
and also the problem of acquiring
the prior knowledge necessary to quickly
solve these tasks (Figure~\ref{firstPageFigure}).
Our solution is an algorithm that learns a DSL
while jointly training a neural network to help write programs in the learned DSL.
\begin{figure}[h]
  \begin{minipage}[c]{0.68\textwidth}
\hspace{-0.5cm}      \begin{tabular}{r|c|}\cmidrule{2-2}
  \sc{Task}&
  \begin{tabular}{rcl}
  Nancy FreeHafer&$\longrightarrow$& Dr. Nancy\\
% Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
 Jane Goodall&$\longrightarrow$&???
  \end{tabular}\\\cmidrule{2-2}
  \sc{Program}&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_1$\code{ s ' '))}\\\cmidrule{2-2}
    \sc{DSL}
&\begin{tabular}{l}
    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b ($\lambda$ (x y) (cons x y)))}\\
    \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s ($\lambda$ (x a)}\\
    \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}\\\cmidrule{2-2}
  \end{tabular}
  \end{minipage}%
  \begin{minipage}[c]{0.31\textwidth}
      \caption{\textbf{Task}: Few-shot learning problem. Model solves tasks by writing \textbf{program}s, and jointly learns a library of reusable subroutines that are shared across multiple tasks, called a \textbf{Domain Specific Language (DSL)}. Program writing is guided by a neural network trained jointly with the library.}\label{firstPageFigure}
    \end{minipage}
\end{figure}

@@

We apply \system to four domains:
list processing; FlashFill-style~\cite{gulwani2011automating} string editing; and symbolic regression.
 For each of these we initially provide a generic
 set of programming primitives.
Our algorithm then discovers
its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
%% \begin{figure}
%%   \begin{tabular}{lcl}
%%     \toprule
%%     Domain&Example Task&Part of the learned DSL\\\midrule
%%     Lists&
%%     \begin{tabular}{rcl}
%%       \code{[7\, 0\, 2]}&$\longrightarrow$&\code{[7\, 0\, 2\, 4]} \\
%%       \code{[3\, 9]}&$\longrightarrow$&\code{[3\, 9\, 4]}
%%     \end{tabular}
%%     &
%%     \begin{tabular}{l}
%%       \code{(foldr nil ($\lambda$ (a b) (cons a b)))}\\
%%           \hspace{0.5cm} \emph{(appends lists)}
%%       \end{tabular}

%%     \\    Strings&
%%     \begin{tabular}{rcl}
%%       Temple Anna H& $\longrightarrow$& TAH\\
%%       Lara Gregori&$\longrightarrow$&LG
%%       \end{tabular}
%%     &
%%     \begin{tabular}{l}
%%       \code{(map ($\lambda$ (x) (if (= x a) b x)))}\\
%%       \hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
%%       \end{tabular}

%%     \\Regression&\includegraphics[width = 1cm]{figures/functions/476.png}&\begin{tabular}{l}
%%       \code{(+ (* real x) real)} \\
%%           \hspace{0.5cm} \emph{(a linear function of x)}
%%       \end{tabular}

%%     \\\bottomrule
%%   \end{tabular}
%%   \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
%% %  \vspace{-4em}
%% \end{figure}

\begin{figure}[hb]
\makebox[\textwidth][c]{
\scriptsize
\renewcommand\code\texttt
\renewcommand\codechar[1]{\texttt{"#1"}}
\begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}}\toprule
  {\normalsize Lists}&{\normalsize Strings}&{\normalsize Symbolic Regression}\\\midrule
  \begin{tabular}{c}
      % \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
      % \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
      % \code{$f(x) = $}\code{($f_0$ nil $x$)}\\\\
      \code{[99\, 4\, 2]}$\to $\code{99}\\
      \code{[2\, 1\, 4\, 6]}$\to $\code{6}\\
      \code{[8\, 7\, 7\, 7]}$\to $\code{8}\\
      \code{$f(\ell) = $}\code{($f_0$ $\ell$)}\\\\
      \code{[7\, 3\, 14\, 6\, 9]}$\to $\code{false}\\
      \code{[9\, 4\, 3\, 4]}$\to $\code{true}\\
      \code{$f(\ell) = $}\code{($f_1$ $\ell$ k)}
  \end{tabular}&
  \begin{tabular}{c}
    +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
    +83 973-831$\to $83.973.831\\
    $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} }\\
    \hspace{1cm}\code{($f_0$ \codechar{.} \codechar{ } (cdr s)))}\\\\
Temple Anna H $\to $TAH\\
Lara Gregori$\to $LG\\
$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}\\

@@

  \begin{tabular}{cc}
    \includegraphics[width = 3em]{figures/functions/4.png}&
    \includegraphics[width = 3em]{figures/functions/146}\\
    \code{$f($x$) = $($f_1$ x)}&    \code{$f($x$) = $($f_6$ x)}\\
    \includegraphics[width = 3em]{figures/functions/112.png}&
        \includegraphics[width = 3em]{figures/functions/92.png}
    \\
    \code{$f($x$) = $($f_4$ x)}&    \code{$f($x$) = $($f_3$ x)}\\

@@

  \begin{tabular}{l}
    % $f_0($\code{r}$,\ell) \,=\, $\code{(fold $\ell$ r ($\lambda$ (x a)}\\
    % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(fold}}\code{(cons (index (length a) $\ell$) a)))}\\
      % \hspace{0.5cm}($f_1$: \emph{Get the largest number})\\
    $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\\hspace{0.8cm}\code{ (if (> a x) a x)))}\\
      \hspace{0.5cm}($f_1$: \emph{Get the largest number in $\ell$})\\
    $f_1(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (empty$?$ $\ell$)}\\
    \phantom{$f_1(\ell,$\code{k}}
    \code{($\lambda$ (x a) (if a a (= k x))))}\\
      \hspace{0.5cm}($f_1$: \emph{Whether $\ell$ contains $k$})\\
  \end{tabular}&

  \begin{tabular}{l}
          $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x)}\\\hspace{1cm}\code{ (if (= x a) b x)) s)}\\
      \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a)}\\\hspace{1.1cm}\code{ (cdr (if (= c x) s a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s empty$?$ car }\\\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}\\
        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})\\
    \end{tabular}&

  \begin{tabular}{l}
    $f_0($\code{x}$)\,=\,$\code{(+ x real)}\\
  $f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} \\
    $f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}\\
    $f_3($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_2$\code{ x)))}\\
    $f_4($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_3$\code{ x)))}\\
    \hspace{0.5cm}\emph{($f_4$: 4th order polynomial)}\\
        $f_5($\code{x}$)\,=\,$\code{(/ real x)}\\
    $f_6($\code{x}$)\,=\,$\code{($f_4$ ($f_0$ x))} \\
    \hspace{0.5cm}\emph{($f_6$: rational function)}\\

@@

\end{tabular}}
\caption{Top: Example tasks from each domain, each followed by the programs \system discovers for them. Bottom: Subset of learned DSL. Notice that learned DSL primitives can call each other ($f_1$/$f_2$ in Strings \& Regression)}
\end{figure}
